Technical Interview

### 2017-09-01 tex/paper/paper.tex
paper: updated all until the Automated testing

\newlength\drop
\makeatletter
\newcommand*\titleM{\begingroup%
\setlength\drop{0.08\textheight}
\centering
\vspace*{\drop}
{\Huge\bfseries Technical Interview}\\[\baselineskip]
%{\scshape MAN-AHL}\\[\baselineskip]
\vfill
\includegraphics[width=0.6\textwidth]{img/logo.png}\par\vspace{1cm}
\vfill
{\Huge\scshape Fadil Mokhchane \\ 
\small \href{https://www.linkedin.com/in/fadil/}{LinkedIn} }\par
\vspace{1cm}
\includegraphics[width=0.2\textwidth]{img/harrington.jpg}\par
\vfill
{\scshape \@date}\par
\vspace*{2\drop}
{\scshape \small Version 0.1}\par
\endgroup}
\makeatother
%% END Memoir customization

@@

\section{The Problem}
We interpret the question as the following the mathematical problem. 
We define a $X$ as a discrete random variable. 
As input, we have a $k$ values, $x_i$, with their associated probabilities, $P_i$.

A discrete probability density function is a probability distribution charaterised 
by a probability mass function :
\[
	\forall i \in \left[1 ...  k\right], P\left[ X = x_i \right] = P_i  
\]
With
\[ 
	\sum_{i = 1}^k  P\left[ X = x_i \right]  = 1
\]
This probability mass function can be illustrated \autoref{fig:distribution} :
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/distribution.png}
\caption{Probability Mass Function}
\label{fig:distribution}
\end{center}
\end{figure}

@@

This way we can build a discrete cumulative function for $X$, $F_x$, such as :
\begin{equation}
	F_x = P\left( X \leq x \right)
	\label{eq:cdf}
\end{equation}
Which can be illustrated with \autoref{fig:cdf2} :
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/cdf.png}
\caption{Discrete Cumulative Distribution Function}
\label{fig:cdf2}
\end{center}
\end{figure}

We then use a normal distribution to generate a number between $0$ and $1$.
\[
	f_x = \frac{1}{\sqrt{2 \pi}} e^{\frac{x^2}{2}}
\]
Using our discrete cumulative function, in \autoref{eq:cdf}, we can pick an $x_i$.
The probability density of the normal distribution with mean $0$ and variance $1$ is:

@@

The key requirement is to deliver a production quality software.
This means that the code has to be well designed, efficient, portable and well tested.

The key features we would like to implement are:
\begin{itemize}
	\item Implement the generator,
	\item Create an User Interface, via Command Line Input,
	\item Compare various pseudo random number generator,
	\item Target $100\%$ testing coverage.
	\item Document the work.
\end{itemize}

\subsection{Design}
Hence we target the following archictecture, c.f. \autoref{fig:architecture1}.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{img/architecture1.png}
\caption{Algorithm App Architecture}
\label{fig:architecture1}
\end{center}
\end{figure}
The various modules implement all the functionalities we need.
\begin{description}
	\item[Analytics] The mathematical functions required,
	\item[Engine] 	The various random numbers generator,
	\item[CLI] 		The command line input module,
	\item[Testing] 	The testing module. 
\end{description}

\subsection{Testing}
We need to test each module independtly, i.e. unit testing, and when they 
are all used together, i.e. integration testing.

Due to our limited testing environment, c.f. \autoref{apx:testing.env},
we limit ourselves to the following tests.

\emph{TBC}

\subsection{UI}
The UI is command line input allowing the user to run each feature.

@@

\section{Solution Proposed}
We can solve the problem by running a query like:
\begin{lstlisting}[
           language=SQL,
           showspaces=false,
           basicstyle=\ttfamily,
           numbers=left,
           numberstyle=\tiny,
           commentstyle=\color{gray}
        ]
SELECT p.name
FROM  (
   SELECT product_id, name FROM product 
      WHERE available_from >=  (today - 1 month)
) p
INNER JOIN (
   SELECT product_id, SUM(quantity) nb FROM orders
      GROUP BY product_id
      WHERE dispatch_date <= (today - 1 year)
) o
   ON p.product_id = o.product_id
WHERE 
   o.nb < 10
\end{lstlisting}

The key issue here in order to produce a production-like quality is 
to test the query proposed.

\subsection{Overview}
We will extend our App by adding a database layer to create and query 
the tables.
We extend also our testing layer to create datasets in order to test
the accuracy and efficiency of our query.

\subsection{Design}
Hence we target the following archictecture, c.f. \autoref{fig:architecture2}.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{img/architecture2.png}
\caption{SQL App Architecture}
\label{fig:architecture2}
\end{center}
\end{figure}

\subsection{Testing}
We unfortunately lack the production tools to analyse the query execution.
Without the access to the query plan of execution, we can only on our
experience and empirical tests to create the best query possible.

Therefore we have simple datasets to test the limiting cases and 
large datasets to test the performance of our query.

We have access to a limited hardware, c.f. \autoref{apx:testing.env}, 
therefore the performance are slow.

\subsection{UI}
We extend the CLI to include the following command :
\begin{description}
	\item[TBC] ...
\end{description}

%-------------------------------------------------------------------------------
\newpage
\chapter{Discussion}
\section{Implementation}
In order to understand and share the creation process, 
we share the Git log in \autoref{apx:git.log}. 
Having a git log allows also rolling back changes, if needed.

We tried to follow a Test Driven Development process as 
demonstrated by the Git log.

\section{Testing}
The testing environment is described in \autoref{apx:testing.env}.

%-------------------------------------------------------------------------------
\newpage
\appendix
\addcontentsline{toc}{chapter}{Appendices}
\chapter{Installation} \label{apx:installation}

Here are the steps to install the App.
\begin{enumerate}
	\item Download Haskell via this \href{https://www.haskell.org/platform/}{link}.
	\item Install following the instructions given by the installer.
	\item Use cabal to download the require packages :
\begin{lstlisting}[
           language=SQL,
           showspaces=false,
           basicstyle=\ttfamily,
           numbers=left,
           numberstyle=\tiny,
           commentstyle=\color{gray}
        ]
	cabal install TBC
\end{lstlisting}
\end{enumerate}

%-------------------------------------------------------------------------------
\newpage
\chapter{Testing Environment} \label{apx:testing.env}

Given that I don't have access to my server this week, 
we run the test on a MacBook Pro 2008 4GB DDR3 Intel Core Duo 2.53GHz.
It has 250GB harddisk, not SSD.

We are using Haskell version 8.2.1 on MacOS.

%-------------------------------------------------------------------------------
\newpage
\chapter{Haskell Libraries} \label{apx:haskell.libraries}

%-------------------------------------------------------------------------------
\newpage
\chapter{Git Log} \label{apx:git.log}
We provide below the git log, as we are simulating a production environment.

%-------------------------------------------------------------------------------
\newpage
\chapter{Code} \label{apx:code}
We provide below the code.

\end{document}

>>>>

\newlength\drop
\makeatletter
\newcommand*\titleM{\begingroup%
\setlength\drop{0.08\textheight}
\centering
\vspace*{\drop}
{\Huge\bfseries Technical Interview}\\[\baselineskip]
%{\scshape MAN-AHL}\\[\baselineskip]
\vfill
\includegraphics[width=0.6\textwidth]{img/logo.png}\par\vspace{1cm}
\vfill
{\Huge\scshape Fadil Mokhchane \\ 
\small \href{https://www.linkedin.com/in/fadil/}{LinkedIn} }\par
\vspace{1cm}
\includegraphics[width=0.3\textwidth]{img/harrington.jpg}\par
\vfill
\vspace*{2\drop}
{\scshape September 1, 2017}\par
{\scshape \small Version 1.0}\par
\endgroup}
\makeatother
%% END Memoir customization

@@

\section{The Problem}
We interpret the question as the following the mathematical problem. 
We define a $X$ as a discrete random variable. 
As input, we have a $k$ values, $x_i$, with their associated probabilities, $p_i$.

A discrete probability density function is a probability distribution charaterised 
by a probability mass function, c.f. \autoref{eq:pdf}, referenced as PDF.
\begin{equation}
	\forall i \in \left[1 ...  k\right], P\left[ X = x_i \right] = p_i  
	\label{eq:pdf}
\end{equation}
With
\[ 
	\sum_{i = 1}^k  P\left[ X = x_i \right]  = 1
\]
The values $\left(x_i, p_i\right)$ are called the PDF pillars.
This probability mass function can be illustrated by \autoref{fig:distribution} :
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/distribution.png}
\caption{Probability Mass Function}
\label{fig:distribution}
\end{center}
\end{figure}

@@

This way we can build a discrete cumulative function, referenced as CDF,
 for $X$, $F_x$, such as :
\begin{equation}
	F_x = P\left( X \leq x \right)
	\label{eq:cdf}
\end{equation}
Which can be illustrated with \autoref{fig:cdf2} :
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/cdf.png}
\caption{Discrete Cumulative Distribution Function}
\label{fig:cdf2}
\end{center}
\end{figure}

We then use a bounded uniform distribution, $U\left(0,1\right)$,
to generate random numbers between $0$ and $1$.
We picked two pseudo random numbers generator : Mersenne Twister\footnote{
The Mersenne twister is a pseudorandom number generator developed by 
Makoto Matsumoto and Takuji Nishimura that is based on a matrix linear 
recurrence over a finite binary field. It provides for fast generation of 
very high quality pseudorandom numbers. 
It has a period of $2^63-1$.
The source for the C code can be found 
\href{http://www.math.sci.hiroshima-u.ac.jp/~m-mat/MT/emt64.html}{here}.
} and Ecuyer\footnote{
This is the default random number generator in haskell one in haskell.
 This implementation uses the Portable Combined Generator of L'Ecuyer
 for 32-bit computers, transliterated by Lennart Augustsson. 
 It has a period of roughly $2.30584e18$.
 This implementation claims to be faster than Ecuyer's one.
 } in order to compare and pick the best one.

Finally using the inverse of the discrete cumulative function, reference as Inverse CDF, 
we can pick an $x_i$, described in \autoref{eq:invcdf}.
\begin{equation}
	\forall y \in \left[0...1 \right], F^{-1}_y = x_i
	\label{eq:invcdf}
\end{equation}
This process gives us a weighted probability pseudo random number generator.

@@

The key requirement is to deliver a production quality software.
This means that the code has to be well designed, simple, 
efficient, portable and well tested.

The key features we would like to achieve :
\begin{itemize}
	\item Implement the weighted probability engine based
	on a uniform random probability generator,
	\item Create an User Interface, via Command Line Input,
	\item Assess the quality of our weighted probability engine,
	\item Compare various pseudo random number generator,
	\item Target $100\%$ testing coverage.
	\item Document the work.
\end{itemize}

\subsection{Design}
We target the following archictecture, c.f. \autoref{fig:architecture1}.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{img/architecture1.png}
\caption{Algorithm App Architecture}
\label{fig:architecture1}
\end{center}
\end{figure}
The various modules implement all the functionalities we need.
\begin{description}
	\item[Analytics] The mathematical functions required,

	In this module we have the implementation of the discrete 
	probability function\footnote{
	We can take as input incomplete probabilities, i.e. that
	don't sum up to $1$. We therefore assign the remaining
	probability to the event we don't select any element.
	Haskell provides the \emph{Maybe a} type to modelise this.
	}, the discrete cumulative function, 
	the inverse discrete cumulative function and the mean and
	standard deviation functions.

	\item[Random]	The uniform pseudo random numbers generator,

	This module contains an abstraction to a uniform pseudo
	random number generator with values within $ \left[ 0, 1 \right]$.
	It can compute a random number or a series of random 
	number and the statistics related to the distribution of 
	the random number generated.
	\item[Engine] 	The weighted probability engine,

	This module includes an abstraction to the weighted 
	pseudo random numbers generators.
	It allows computation and statistics on these numbers.
	It inluce the requested nextNum function.
	\item[CLI] 		The command line input layer,

	As it is meaningless to look only at one random number at the time.
	The CLI provides a flexible API to compute $n$ random numbers.
	We can select any random number we have implemented.
	We display only the distribution of these random numbers.
	\item[Testing] 	The testing module,

	This module provides all the automated tests. They are unit 
	tests at the module level and integration tests at the key
	features level.
\end{description}

\subsection{UI}
The UI is a command line input allowing the user to run each feature.
After building the project, with the following command:
\begin{verbatim}
cabal build
\end{verbatim}
The binary is located in \emph{./dist/build/MAN-AHL/MAN-AHL}.
We can run weighted engine or the unform generator
 as described in \autoref{fig:cli}, with the following command:
\begin{verbatim}
time ./dist/build/MAN-AHL/MAN-AHL -run=Weighted -nSims=1000000
\end{verbatim}
 The binary has the option \emph{-help}, stating all the default values.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/cli.png}
\caption{Command Line Input}
\label{fig:cli}
\end{center}
\end{figure}

\section{Implementation}

\section{Automated Unit and Integration Testing}
We need to test each module independtly, i.e. unit testing, and when they 
are all used together, i.e. integration testing.

The testing environment used is described in below\footnote{
Given that I don't have access to my server, 
we run the tests on a MacBook Pro 2008 4GB DDR3 Intel Core Duo 2.53GHz.
It has 250GB harddisk, not SSD. 
We are using Haskell version 8.2.1 on MacOS. Vim version 8.0.
}.

We use heavily a random testing library, i.e. QuickCheck\footnote{
QuickCheck is a library for random testing of program properties.
The API is described 
\href{http://hackage.haskell.org/package/QuickCheck-2.10.0.1/docs/Test-QuickCheck.html}{here}.
}, to test comprehensively the software.
Automated pseudo random tests allowed to discover issues and
test cases that were not included at first.

Our test data is mostly pseudo random: 
PDF pillars, PDF, CDF and Inverse CDF inputs.

We can run the automated as described in  \autoref{fig:tests} with the command:
\begin{verbatim}
time cabal test --show-details=streaming
\end{verbatim}
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/tests.png}
\caption{Automated Testing}
\label{fig:tests}
\end{center}
\end{figure}

The following Test Suites, i.e. lists of tests, compose our test plan split by modules.

\subsection{Test Data}

The key sample used throughout the testing is the generation of PDF pillars.
We tried to pick a realistic sample.

The pillars can have 1 to 10 pillars. Their summed probability are always less 
than or equal to 1.  Each probability is within the $[0.01, 1]$ bound.

\subsection{Analytics Testing} 
These are the unit tests for the Analytics only.
\begin{description}
	\item [CDF creation]
		Given a 1000 pseudo random PDF, we ensure 
		that the last pillar of the CDF is 1.
		As the CDF is build incrementally, this gives a good sense that
		the CDF construction has been correct.
	\item [CDF and PDF pillars check]
		Given a 1000 pseudo random PDF, we ensure that
		the pillars on the x axis of the PDF are the pillars on the y
		axis of the CDF.
	\item [PDF Stable]
		Given a 1000 pseudo random PDF pillars, 
		we ensure that we get the same PDF given the reverse input.
	\item [PDF consistency]
		Given a 1000 pseudo random PDF, 
		we ensure that the sum of the PDF pillars probabilities are 
		less than 1. We indeed modelise the non selected element
		only in the CDF.
	\item [Inverse CDF valid]
		Given a 1000 pseudo random PDF,
		we build the CDF and we ensure that for a random number of
		x values within 0 and 1 we always recover a x pillars of the PDF.
	\item [PDF Failure]
		We test the expecture failure of the PDF construction. 
		We can't have negative values. We need at least one valid pillar.
		The pillars with 0 probabilities are discarded.
		And the sum of the PDF pillars probabilities must be less or
		equal to one.
	\item [CDF Failure]
		 The CDF construction always assume a correct PDF as
		 input. If it is not the case, it fails if the aggregated 
		 probabilities are greater than 1.
		 The test ensure an exception is raised.
	\item [Inverse CDF Failure]
		We check that the inverse CDF function raise an
		exception if the x input is outside the correct bounds, 
		$\left[0, 1\right]$.
	\item [Mean and Standard Deviation Failure]
		These functions raise an exception for empty lists.
		We test it automatically.
	\item [PDF and CDF Non Regression]
		We ensure without randomness that given a set of pillars,
		we recover the expected CDF and PDF.
\end{description}
\subsection{Random Testing}

These are the unit tests for the uniform pseudo random generator.

\begin{description}
	\item [Mean and Standard Deviation]
		Given a large number of random number generated for
		both Ecuyer and Mersenne, i.e. $100000$.
		We ensure that the mean and the standard deviation 
		of the distributions are still the same. Run 100 times.
	\item [Bounds]
		Test ensuring that the uniform pseudo random generators
		generate probabilities with the $[0, 1]$ bound.
		Run $100$ times for $100000$ random numbers.
	\item [Uniform]
		Ensure that the ditributions generated are uniform.
		Run 10 times for $100000$ random numbers.
	\item [Performance]
		Ensure that the pseudo random numbers keep at
		least the same performance.
		Run 10 times for $100000$ random numbers.
\end{description}

\subsection{Engine Testing} 
These are the unit tests for the weighted pseudo random engine.
All tests are run using the two available uniform pseudo random
generators. 
These are like the integration tests as they use most of the library.
\begin{description}
	\item [Distribution check]
		We check that we recover the PDF distribution from
		a sample of 1000000 at $0.5\%$ precision.
		We run this for 10 pseudo random PDF.
	\item [Engine Fail]
		Ensure that the construction of the engine 
		handles all the expected failure.
	\item [Performance]
		Ensure that the engine keep at least the same performance.
		Run 10 times for $100000$ random  numbers	
\end{description}

\subsection{CLI Testing}
We test that the API produce what is expected.
\begin{description}
	\item [Query]
	We test that given an argument list, we produce the expected
	Query.
	\item[Run]
	We test that for given a Query, we produce the expected Result.
\end{description}

\section{Qualitative Testing}

\section{Performance Testing}

@@

\section{Solution Proposed}
We can solve the problem by running a query like:
\begin{lstlisting}[
           language=SQL,
           showspaces=false,
           basicstyle=\ttfamily,
           numbers=left,
           numberstyle=\tiny,
           commentstyle=\color{gray}
        ]
SELECT p.name
FROM  (
   SELECT product_id, name FROM Product 
      WHERE available_from <=  (today - 1 month)
) p
LEFT OUTER JOIN (
   SELECT product_id, SUM(quantity) nb FROM Orders
      WHERE dispatch_date >= (today - 1 year)
      GROUP BY product_id
      HAVING nb < 10
) o
   ON p.product_id = o.product_id
\end{lstlisting}

Adding indexes to the columns on which we join and maybe query would improve 
the performance. 
A query plan of the execution would indicate what is the best design for this 
database scheme.

%-------------------------------------------------------------------------------
\newpage
\appendix
\addcontentsline{toc}{chapter}{Appendices}

\chapter{Code} \label{apx:code}
We provide below the code.

%-------------------------------------------------------------------------------
\newpage
\chapter{Installation} \label{apx:installation}

Here are the steps to install the App.
\begin{enumerate}
	\item Download Haskell via this \href{https://www.haskell.org/platform/}{link}.
	\item Install following the instructions given by the installer.
	\item Use cabal to download the require packages :
\begin{lstlisting}[
           language=SQL,
           showspaces=false,
           basicstyle=\ttfamily,
           numbers=left,
           numberstyle=\tiny,
           commentstyle=\color{gray}
        ]
	cabal install TBC
\end{lstlisting}
\end{enumerate}
\end{document}
